



```{r warning = FALSE, message = FALSE}
# Suppress dplyr summarise grouping warning messages
options(dplyr.summarise.inform = FALSE)

library(tidyverse)
library(tidymodels)
library(ggplot2)
credit_card_df <- readRDS(url('https://gmubusinessanalytics.netlify.app/data/credit_card_df.rds'))

```

```{r warning = FALSE, message = FALSE}
credit_card_df
```


# Data Analysis

**Question**:

Do spend and transaction ratios have any effect on customer status?

**Answer**:

Most of the customers who have spent more in Q1 than in Q4 and also those with greater transactions in Q1 compared to Q4 have closed their accounts.

```{r}

ggplot(credit_card_df, aes(x=transaction_ratio_q4_q1,y=spend_ratio_q4_q1))+geom_point(aes(col=customer_status))+xlab('Transaction ratio between Q4 and Q1 quarters')+ylab('Spend ratio between Q4 and Q1 quarters')+ggtitle('Effect of Spend and transaction ratios on Customer Status')+theme_light()


```

**Question**:

Does particular card type has many closed accounts?

**Answer**:

customers with blue card have high probability of closing their accounts than those with other cards.

```{r}

credit_card_df %>% filter(customer_status=='closed_account') %>% group_by(card_type) %>% summarise(Closed_account_Count=n()) %>% mutate(Closed_Percentage=(Closed_account_Count/sum(Closed_account_Count))*100) %>% select(card_type,Closed_Percentage) %>% arrange(desc(Closed_Percentage))

```


**Question**:

Does the number of months inactive has anything to do with customers closing their account? On an average,how long customers are loyal to the bank?

**Answer**:

Generally, customers who are highly inactive are most likely to close their accounts but this data shows high inactivity has nothing to do with customers closing their account and the average time a customer stays before closing his/her account is around 36 months.


```{r}
credit_card_df %>% filter(customer_status=='closed_account') %>% group_by(months_inactive_last_year) %>% summarise(Closed_account_Count=n()) %>% arrange(desc(Closed_account_Count))

credit_card_df %>% filter(customer_status=='closed_account') %>% summarise(Avg_months_before_closing=mean(months_since_first_account))

```

**Question**:

what is the influence of employment on the customers closing their accounts?

**Answer**:
Customers with part-time jobs have high probability of closing their accounts followed by full-time employed and self-employed customers respectively.


```{r}


credit_card_df %>%
  group_by(employment_status, customer_status) %>%
  summarise(customer_count = n()) %>% mutate(pct=customer_count/sum(customer_count)) %>%
ggplot(aes(x="", y=pct, fill=customer_status)) +
geom_bar(stat="identity", width=1) +
coord_polar("y", start=0) + facet_wrap(~employment_status)+
geom_text(aes(label = paste0(round(pct*100), "%")), position = position_stack(vjust = 0.5))+
scale_fill_manual(values=c('red3','greenyellow')) +
ggtitle("Influence of Employment on customer status") + theme_void() 




```


**Question**:

Customers of which education level are mainly closing their accounts?

**Answer**:

Majority of the customers who closed their accounts are doctorate degree holders and most of those with bachelors degree are still continuing their accounts in the banks. 

```{r}
credit_card_df %>% group_by(education,customer_status) %>% summarise(customer_Count=n()) %>% mutate(per=(customer_Count/sum(customer_Count))*100) %>%
ggplot(aes(x=education,y=per,fill=customer_status))+geom_bar(stat = 'identity',position = 'dodge')+ geom_text(aes(label = paste0(round(per), "%")),vjust=1.2, position = position_dodge(0.8))+xlab('Education Level')+ylab('Percentage')+ggtitle("Impact of Education level on customer status")+theme(plot.title = element_text(size=12,hjust=0.5,vjust = 0.5))+scale_fill_manual(values=c('firebrick3','palegreen1'))

```

**Question**:

What are the average credit card limits and incomes for closed and active accounts?

**Answer**:
Average credit card limit and average income are on higher side for active account holders though they differ slightly.

```{r}
credit_card_df %>%
  group_by(customer_status) %>%
  summarise(Avg_cred_limit =mean(credit_limit))

credit_card_df %>%
  group_by(customer_status) %>%
  summarise(Avg_income =mean(income))


```


**Question**:
How is closed accounts spread across different age groups?

**Answer**:
Middle aged customers are more likely to close their accounts.

```{r}
credit_card_df %>% filter(customer_status=='closed_account') %>% ggplot(aes(x=age))+
  geom_density(fill='lightsteelblue1')+
  xlab("Age")+
  ggtitle("Distribution of Age for closed accounts")+
  theme_light()+
  theme(legend.position = "bottom")+theme(plot.title = element_text(size=12,hjust=0.5,vjust = 0.5))
```

# Machine Learning

```{r}
#Initial proportion of Active and Closed accounts

credit_card_df %>% group_by(customer_status) %>% summarise(cnt=n()) %>% mutate(prop=cnt/sum(cnt)) %>% select(customer_status,prop)


#Data Splitting
set.seed(367821)
customer_split <- initial_split(credit_card_df, prop = 0.75, 
                                strata = customer_status)

customer_training = customer_split %>% training()

customer_test = customer_split %>% testing()

set.seed(17897)

customer_folds <- vfold_cv(customer_training, v = 5)

my_metrics <- metric_set(accuracy, roc_auc)


```

```{r}
#Feature Engineering

customer_recipe <- recipe(customer_status ~ ., data = customer_training) %>% 
                       step_YeoJohnson(all_numeric(), -all_outcomes()) %>% 
                       step_normalize(all_numeric(), -all_outcomes()) %>% 
                       step_dummy(all_nominal(), -all_outcomes())


customer_recipe %>% 
  prep(training = customer_training) %>% 
  bake(new_data = NULL)

```


# Logistic Model

```{r}
# logistic regression
customer_logistic <- logistic_reg() %>% 
                  set_engine('glm') %>% 
                  set_mode('classification')

# workflow
logistic_wf = workflow() %>%
  add_model(customer_logistic) %>% add_recipe(customer_recipe)

#model fit
log_last_fit <- logistic_wf %>% 
                  last_fit(split = customer_split,
                           metrics = my_metrics)

metrics = log_last_fit %>% 
  collect_metrics()

metrics

log_last_fit_results <- log_last_fit %>% 
                     collect_predictions()

log_last_fit_results

#Roc curve
log_last_fit_results %>% 
  roc_curve(truth = customer_status, estimate = .pred_closed_account) %>% 
  autoplot()

#Confusion matrix
conf_mat(log_last_fit_results, truth = customer_status, estimate = .pred_class)

```

# KNN Model

```{r}
customer_knn <- nearest_neighbor(neighbors = tune()) %>% 
  set_engine('kknn') %>% 
  set_mode('classification')

knn_wf <- workflow() %>% 
  add_model(customer_knn) %>% 
  add_recipe(customer_recipe)

knn_grid <- tibble(neighbors = c(12, 25, 40, 60, 80, 120))

## Tune  workflow
set.seed(7698)

knn_tuning = knn_wf %>% 
  tune_grid(resamples = customer_folds,
            grid = knn_grid)

knn_tuning %>% collect_metrics()

best_k = knn_tuning %>% select_best(metric = 'roc_auc')

final_knn_wf <- knn_wf %>% finalize_workflow(best_k)

knn_last_fit <- final_knn_wf %>% last_fit(split = customer_split,metrics=my_metrics)

knn_last_fit %>% collect_metrics()

knn_predictions <- knn_last_fit %>% collect_predictions()

knn_predictions

#Roc curve
knn_predictions %>% 
  roc_curve(truth = customer_status, estimate = .pred_closed_account) %>% 
  autoplot()

#confusion matrix
conf_mat(knn_predictions, truth = customer_status, estimate = .pred_class)


```


# Random forest Model

```{r}

library(vip)
customer_rf<- rand_forest(mtry = tune(),
                        trees = tune(),
                        min_n = tune()) %>% 
            set_engine('ranger', importance = "impurity") %>% 
            set_mode('classification')

rf_wf <- workflow() %>% 
               add_model(customer_rf) %>% 
               add_recipe(customer_recipe)

set.seed(8796)

rf_grid <- grid_random(mtry() %>% range_set(c(2, round(sqrt(ncol(customer_training))))),
                       trees(),
                       min_n(),
                       size = 7)

set.seed(9786)

rf_tuning <- rf_wf %>% 
             tune_grid(resamples = customer_folds,
                       grid = rf_grid)

best_rf <- rf_tuning %>% 
           select_best(metric = 'roc_auc')

final_rf_workflow <- rf_wf %>% 
                     finalize_workflow(best_rf)

rf_last_fit <- final_rf_workflow %>% 
               last_fit(split = customer_split,metrics=my_metrics)

rf_vip_fit <- rf_last_fit %>% 
          extract_fit_parsnip()

vip(rf_vip_fit)

metrics<-rf_last_fit %>% collect_metrics()

metrics

#Roc curve
rf_last_fit %>% collect_predictions() %>% 
                roc_curve(truth  = customer_status, estimate = .pred_closed_account) %>% 
                autoplot()

#Confusion Matrix
rf_last_fit %>% collect_predictions() %>% conf_mat(truth = customer_status, estimate = .pred_class)


```












